Meta-analysis of the Interoceptive Accuracy Scale (IAS) Structure and its Subjective Correlates

### 2025-03-03 paper/1_Initial/manuscript.qmd
minor

Interoception is referred to the process of sensing, interpreting and integrating information pertaining to internal organs, such as the heart, the lungs or the gut [@khalsa2018interoception].
While recent research emphasizes a key role of interoception in a variety of processes (e.g., emotion regulation, decision making) and of outcomes (physical and psychological well being), the field remains clouded by concerns about how interoception is assessed.

@@

<!-- Various measures of interoception have been developed (see @fig-measures), forming a combination of "objective" and "subjective" assessments (i.e., physiological tasks such as the heart beat counting or tracking *vs.* questionnaires and subjective scales involving a metacognitive reflection), "explicit" and "implicit" paradigms (i.e., directing participants' awareness and attention to interoceptive processes *vs.* measuring interoception unbeknownst to them), various interoceptive modalities (e.g., cardioception, respiroception, gastroception) and theoretical dimensions (e.g., accuracy, sensitivity, awareness). -->

<!-- While this is longer it breaks it down a bit more -->
Various measures of interoception have been developed (see @fig-measures), combining different approaches, including both “objective” and “subjective” assessments—physiological tasks, such as  Heartbeat Counting Task [HCT, @schandry1981heart] or the Heartbeat Detection Task [HDT, @kleckner2015methodological], versus questionnaires and rating scales that require metacognitive reflection. 
A further distinction can be made between “explicit” and “implicit” paradigms, where explicit measures direct participants’ awareness toward interoceptive processes, while implicit measures assess interoception unbeknownst to them. 
Additionally, interoceptive measures vary across different modalities, such as cardioception, respiroception, and gastroception. 
Finally, these measures assess various theoretical dimensions, including accuracy, sensitivity, awareness, and magnitude.
While there is no consensus as to which particular approach provides the most accurate and "pure" measure of interoception and interoceptive abilities (assuming it is a unidimensional construct), it is instead plausible that each measure has strengths and limitations, and a utility dependent on the context and goal at hand [@jahedi2014; @Desmedt2023accuracy].

@@

Although the use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical at first, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information.
These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively [@khalsa2018interoception; @suksasilp2022towards].
Moreover, the notion that self-reports might not reflect the same processes as other interoception tasks might be important to contextualize the apparent lack of convergence between measures in the field [@desmedt2022measures]. 
For instance, existing findings typically show weak or no correlations between questionnaires and objective measures [including for measures of the same theorethical dimensions, e.g., task-based accuracy vs. self-reported accuracy, @murphy2019; @brand2023; @arslanova2022], such as the HCT and the HDT. 
Additionally, objective measures assessing the same interoceptive dimension, such as accuracy, either show no correlation [e.g., @brand2023] or only weak correlations [e.g., @hickman2020relationship].
Perhaps more surprisingly, low correlations have been observed even among questionnaires, suggesting the targeting of different facets related to interoception.

<!-- 
Garfinkel - sensibility-> Self-perceived dispositional tendency to be internally self-focused and interoceptively cognisant,
Khalsa - sensibility -> Self-perceived tendency to focus on interoceptive stimuli (trait measure)
Khalsa self-report scales -> The ability to reflect upon one’s autobiographical experiences of interoceptive states, make judgments about their outcomes, and describe them through verbal or motor responses. -->

One striking example concerns the assessment of interoceptive sensibility, which is broadly defined as the self-reported tendency to focus on and detect internal sensations [@garfinkel2015knowing], but more narrowly as the subjective tendency to focus on interoceptive signals, without necessarily implying detection ability [@khalsa2018interoception].
A recent systematic review suggested that various questionnaires designed to assess interoceptive sensibility may, in fact, measure distinct constructs, leading researchers to treat them as equivalent despite overall low convergence [@desmedt2022measures].

Notably, this review adopted a broad definition of 'sensibility,' incorporating both interoceptive sensibility and interoceptive self-report scales, as described in @khalsa2018interoception's eight-facet model. 
In this framework, self-report measures are categorized into interoceptive sensibility and interoceptive self-report scales (i.e., the ability to evaluate one's personal experiences of internal bodily states, assess their consequences, and articulate them through speech or movement).
Several widely used questionnaires were included in the review, such as the Multidimensional Assessment of Interoceptive Awareness [MAIA, @mehling2012; MAIA-2, @mehling2018multidimensional], the Body Perception Questionnaire [BPQ, @porges1993body], the Private subscale of the Body Consciousness Questionnaire [PBCS, @miller1981consciousness], the Body Awareness Questionnaire [BAQ, @shields1989body], and the Eating Disorder Inventory [EDI, @garner1991eating; @garner1983].
The lack of correlations to moderate correlations among these questionnaires highlight the need for greater conceptual clarity regarding what each measure captures, how they relate to different dimensions of interoception, and their potential overlaps with other constructs, such as alexithymia and body awareness.

@@

A recently developed scale with a rapidly growing popularity is the Interoceptive Accuracy Scale [IAS, @murphy2019].
The IAS consists of 21 Likert-scale items that query how accurately one can perceive different bodily signals, with one item per physiological modality such as respiration (*"I can always accurately perceive when I am breathing fast"*), heart (*"I can always accurately perceive when my heart is beating fast"*), skin (*"I can always accurately perceive when something is going to be ticklish"*), arousal or bodily functions like coughing (*"I can always accurately perceive when I am going to cough"*) or urinating (*"I can always accurately perceive when I need to urinate"*).
Notably, the IAS' statements are about specific interoceptive behaviours, which is a distinct difference with other popular interoception questionnaires, such as the MAIA-2, which contains more general and metacognitive items (e.g., *"I trust my body sensations"*, *"I can notice an unpleasant body sensation without worrying about it"*), as well as dimensions related to attention regulation (e.g., Not-distracting) or emotion regulation (e.g., Not-worrying).

@@

**note to dom review this. 1. yes I think they are general interoceptive signals (I added the items that loaded to both factors for Murphy's and Koike's validation) 2. Do we want to add that lin created a short verison of the scale? 3. Yes kioke  did named their factors**

Although the original validation study suggested a two-factor structure for the IAS, reflecting the perception of general interoceptive signals (i.e., urinate, hungry, defecate, thirsty, pain, heart, taste, breathing, temperature, mulscles, affective touch, vomit, sexual arousal), and that of signals that may be difficult to perceive solely through interoceptive information (i.e., itch, tickle, cough, burp, bruise, blood sugar, sneeze, wind). 
The authors however underlined its acceptable but imperfect fit [@murphy2019, p. 127], and several follow-up studies have indeed identified different optimal solutions. 
For instance, @brand2023 reported a 1-factor solution, while @lin2023 and @campos2021 [the former using a new approach to structure analysis, Exploratory Graph Analysis, @golino2017exploratory] found bifactor solutions - one general factor above a set of lower-level factors [@rodriguez2016evaluating] - to be the best fit.

Discussions have also been focused on specific items.
For instance, @murphy2019 notes that some items might measure direct interoceptive signals such as cardioception, while others might capture phenomena not perceivable through interoceptive signals alone (e.g., "bruising"; p. 119).
@lin2023 also highlights their correlation analysis, showing five locally dependent pairs and three items (touch, blood sugar, bruise) with exceptionally high difficulty and low discrimination, and @campos2021 reported "tickle" to be the only item that reflected more specific factors than the general factor.

Localization issues also arose in @lin2023 validation of the IAS, where both "itch" and "tickle" correspond to the same Chinese character, leading to their collapse into a single item. 
Notably, the only other study to report a 2-factor solution was conducted by @koike2023, who performed an Exploratory Factor Analysis (EFA) assuming 2 factors to align with the findings from the original validation paper. 
According to the authors, the first factor appears to reflect cutaneous sensations (e.g., itching, tickling, coughing, burping, affective touch, bruising, passing gas, sneezing, muscle sensations, sexual arousal, and taste), while the second factor seems to represent visceral sensations (e.g., urination, defecation, hunger, thirst, pain, breathing, fatigue/blood sugar, temperature, vomiting, and heartbeat).

@@

**Note to dom: @ferentzi2021examining suggests that all MAIA dimensions, except for Not-Worrying and Not-Distracting, relate to a general interoceptive awareness factor. The low correlations of NW and ND with this general factor indicate that these dimensions do not contribute significantly to the measurement of general interoceptive awareness. Instead, NW appears to measure low emotionality or neuroticism, while ND does not capture a coherent underlying structure.**

Regarding its validity, the IAS has naturally been compared to other interoception-related measures, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional], except for the Not-Distracting and Not-Worrying subscales [@brand2023] - which were highlighted previously as related to non-interoceptive abilities [@ferentzi2021examining].
Interestingly, findings on the correlation between the IAS and the body awareness dimension of the Body Perception Questionnaire [BPQ-A, @porges1993body] have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019].
Small positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023].
Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [@lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia, @brand2023; @murphy2019].
Lastly, small positive correlations have also been reported with the Interoceptive Attention Scale [IATS, @lin2023, @koike2023], though studies have also found no correlation between these measures [@gabriele2022dissociations].

@@

**TODO: can you update the fig. 1's box (assessmeent of related traits) to remove the question marks and put the the stronger interoception correlates** - as in with what?

>>>>

Interoception is referred to the process of sensing, interpreting and integrating information pertaining to internal organs, such as the heart, the lungs or the gut [@khalsa2018].
While recent research emphasizes a key role of interoception in a variety of processes (e.g., emotion regulation, decision making) and of outcomes (physical and psychological well being), the field remains clouded by concerns about how interoception is assessed.

@@

Various measures of interoception have been developed (see @fig-measures), forming a combination of "objective" and "subjective" assessments (i.e., physiological tasks such as the heart beat counting or tracking *vs.* questionnaires and subjective scales involving a metacognitive reflection), "explicit" and "implicit" paradigms (i.e., directing participants' awareness and attention to interoceptive processes *vs.* measuring interoception unbeknownst to them), various interoceptive modalities (e.g., cardioception, respiroception, gastroception) and theoretical dimensions (e.g., accuracy, sensitivity, awareness).
While there is no consensus as to which particular approach provides the most accurate and "pure" measure of interoception and interoceptive abilities (assuming it is a unidimensional construct), it is instead plausible that each measure has strengths and limitations, and a utility dependent on the context and goal at hand [@jahedi2014].

@@

Although the use of subjective self-report questionnaires to measure deeply embodied functions might seem paradoxical at first, recent redefinitions of interoception emphasize the role of high-level and metacognitive elaboration of interoceptive information.
These redefinitions provide theoretical grounding to support the idea that some facets of interoception, including participants' metacognitive beliefs, can be assessed subjectively [@khalsa2018interoception; @suksasilp2022towards].
Moreover, the notion that self-reports might not reflect the same processes as other interoception tasks might be important to contextualize the apparent lack of convergence between measures in the field [@desmedt2022measures]. For instance, existing findings typically show weak or no correlations between questionnaires and objective measures [including for measures of the same theorethical dimensions, e.g., task-based accuracy vs. self-reported accuracy, @murphy2019; @brand2023], such as the Heartbeat Counting Task [HCT, @schandry1981heart] and the Heartbeat Detection Task [HDT, @kleckner2015methodological]. 
Perhaps more surprisingly, low correlations have been observed even among questionnaires, suggesting the targeting of different facets related to interoception.

One striking example concerns the assessment of "interoceptive sensibility", defined as the self-reported tendency to focus on and detect internal sensations [@garfinkel2015knowing], or more narrowly as the subjective ability to focus on interoceptive signals [@khalsa2018interoception]. **TODO: ANA check if the definitions are correct**.
A recent systematic review suggested that various questionnaires [**WHICH ONES**] designed to assess this dimension actually measure distinct constructs, leading researchers to mistakenly treating them as equivalent measures [@desmedt2022measures].
This underlines the need for a more comprehensive understanding of what is being measured with different questionnaires and dimensions, as well as their potential overlaps with other constructs (e.g., alexithymia, body awareness).

@@

A recently developed scale with a rapidly growing popularity is the Interoceptive Accuracy Scale [IAS, @murphy2019].
The IAS consists of 21 Likert-scale items that query how accurately one can perceive different bodily signals, with one item per physiological modality such as respiration (*"I can always accurately perceive when I am breathing fast"*), heart (*"I can always accurately perceive when my heart is beating fast"*), skin (*"I can always accurately perceive when something is going to be ticklish"*), arousal or bodily functions like coughing (*"I can always accurately perceive when I am going to cough"*) or urinating (*"I can always accurately perceive when I need to urinate"*).
Notably, the IAS' statements are about specific interoceptive behaviours, which is a notable difference with other popular interoception questionnaires, such as the Multidimensional Assessment of Interoceptive Awareness scale [MAIA, @mehling2012; MAIA-2, @mehling2018multidimensional], which contains more general and metacognitive items (e.g., *"I trust my body sensations"*, *"I can notice an unpleasant body sensation without worrying about it"*), as well as dimensions related to attention regulation (e.g., Not-distracting) or emotion regulation (e.g., Not-worrying).

@@

Although the original validation study suggested a two-factor structure for the IAS, reflecting the perception of [**general?**] interoceptive signals, and that of signals that may be difficult to perceive solely through interoceptive information. 
The authors however underlined its acceptable but imperfect fit [@murphy2019, p. 127], and several follow-up studies have indeed identified different optimal solutions. 
For instance, @brand2023 and @golino2017exploratory (the latter using a new approach to structure analysis, Exploratory Graph Analysis), reported a 1-factor solution, while @lin2023 and @campos2021 found bifactor solutions - one general factor above a set of lower-level factors [@rodriguez2016evaluating] - to be the best fit. 
Discussions have also been focused on specific items.
For instance, @murphy2019 notes that some items might measure direct interoceptive signals such as cardioception, while others might capture phenomena not perceivable through interoceptive signals alone (e.g., "bruising"; p. 119).
@lin2023 also highlights their correlation analysis, showing five locally dependent pairs and three items (touch, blood sugar, bruise) with exceptionally high difficulty and low discrimination, and @campos2021 reported "tickle" to be the only item that reflected more specific factors than the general factor.
Localization issues also arisen with two items, **TODO** (talk about chinese stuff)

<!-- Notably, the only other study to report a 2-factor solution was conducted by @koike2023, who performed an Exploratory Factor Analysis (EFA) assuming 2 factors to align with the findings from the original validation paper. -->
**TODO: how did koike2023 interpret the two factors? Do they say?**

@@

Regarding its validity, the IAS has naturally been compared to other interoception-related measures, and shows a positive correlations with most facets of the MAIA [@mehling2018multidimensional], except for the Not-Distracting and Not-Worrying subscales [@brand2023] - which were highlighted previously as related to non-interoceptive abilities [**do we have a ref to say that? will also be useful for our MINT paper to back up our claims that these 2 dims are something else**].
Interestingly, findings on the correlation between the IAS and the body awareness dimension of the Body Perception Questionnaire [BPQ-A, @porges1993body] have been mixed: some studies report small positive correlations [@brand2023; @campos2021; @koike2023], while others find small negative correlations [@lin2023] or no correlation at all [@murphy2019].
Small positive correlations have also been observed with the "observation" and "description" subscales of the Five Facet Mindfulness Questionnaire [FFMQ, @baer2006using; @brand2023; @koike2023], as well as with the "non-reactivity" and "acting with awareness" subscales [@koike2023].
Additionally, the IAS has shown a positive correlation with the interoceptive awareness subscale of the Eating Disorder Inventory [@lin2023] and a negative correlation with the Interoceptive Confusion Questionnaire [ICQ, @brewer2016alexithymia; @brand2023; @murphy2019].
Lastly, small positive correlations have also been reported with the Interoceptive Attention Scale [IATS, @lin2023; @koike2023], though studies have also found no correlation between these measures [@gabriele2022dissociations].
Interestingly, the IAS and the IATS supposedly measure two different interoceptive processes (i.e., accuracy and attention, respectively) which contradict @murphy2019 proposed 2x2 framework. **TODO: how so? Elaborate**

@@

**TODO: can you update the fig. 1's box (assessmeent of related traits) to remove the question marks and put the the stronger interoception correlates**
